### 2018-04-23 manuscript/case.tex
manuscript added

\section{Case Study}

\subsection{OVERVIEW Case 1: City Folding}

compare the distorted maps from various groups

\subsection{DETAIL Case 2: Look into Detail}

check the detail information of a group