### 2018-05-04 manuscript/relatedwork.tex
manuscript updated

\section{Related Work}

The related work is discussed around the related methods in map distortion, and the work analyzing facility accessbility....

\textbf{Anonymous Movement Analysis} Introduce the classifical movement analysis focus. Data mining (Zheng Yu's work), Specially in the field of visual analytics, also most of work focus on the efficiency of the travel infrustions, (Wang, Lu, Zeng, Qu, Wu Yingcai).

\textbf{Social Related Analysis} In the GIS field, (Liu Yu), Soical Media (Chen, CaoNan)