### 2018-05-14 manuscript/relatedwork.tex
intro, data, related updated

\section{Related Work}

The related work is discussed around the related methods in map distortion, and the work analyzing facility accessbility....

\textbf{Anonymous/Conventional Movement Analysis} Introduce the classifical movement analysis focus. Data mining (Zheng Yu's work), Specially in the field of visual analytics, also most of work focus on the efficiency of the travel infrustions, (Wang, Lu, Zeng, Qu, Wu Yingcai).

\textbf{Social Related Analysis} As the mass adaption of social medial serivces, such as Twitter, Facebook, ti has become an hot research area of ... The new and rich data as Twitter flowerish the research. ... As social media data continues attracting the attention from research fields, such social science, there are lot of research performed around this topics. Most of work infer key characteristics from individuals using social media.

In the GIS field, (Liu Yu), Soical Media (Chen, CaoNan)

\cite{Longley2015}\cite{Paul2016_twitter} identifies and assess the biases inherent in social media usage in social research and evaluate the deployment of social media data in research applications.