### 2018-05-15 manuscript/relatedwork.tex
rw updated

\section{Related Work}

The related work is discussed around the related methods in map distortion, and the work analyzing facility accessbility....

As the involvement of various data collected in urban context, the unpercedent opportunities give rise to the emerge of urban computing(zhengyu). It also triggers a broad use of visual analytics, to do data exploration and pattern interpretation, and visual learning~\cite{zheng2016visual}.

\textbf{Anonymous/Conventional Movement Analysis} (Introduce the conventional traffic analysis) Introduce the classifical movement analysis focus. Data mining (Zheng Yu's work), Specially in the field of visual analytics, also most of work focus on the efficiency of the travel infrustions, Such as traffic jam propagation~\cite{wang2013visual}, route travel behaviour~\cite{lu2015trajrank} (Wang, Lu, Zeng, Qu, Wu Yingcai. zeng wei). ~\cite{andrienko2010discovering}.

\textbf{Demographical Analysis} As the mass adaption of social medial serivces, such as Twitter, Facebook, ti has become an hot research area of ... The new and rich data as Twitter flowerish the research. ... As social media data continues attracting the attention from research fields, such social science, there are lot of research performed around this topics. Most of work infer key characteristics from individuals using social media. Liu et al.~\cite{liu2015social} emphasize the geospatial big data as social sensing as an analogue of remote sensing data in social science research.

Social media...
In the GIS field, (Liu Yu), Chen et al.~\cite{chen2016interactive} presents a visual analytics system to support the exploraiton of sparcely sampled trajectory from social media. Wu et al. also explore the ~\cite{wu2016telcovis}

(Chen, CaoNan)

Although real information is not enforced, a significant fractions of users tend to provide real names~\cite{peddinti2014internet}, which make it possible to breakdown the demographic characteristics of social media users. Luo et al.~\cite{luo2016explore} derive race, gender and age as three demographics dimensions to analyze its impact on the urban human mobility patterns. 
\cite{Longley2015}\cite{Paul2016_twitter} identifies and assess the biases inherent in social media usage in social research and evaluate the deployment of social media data in research applications.