### 2018-05-22 manuscript/case.tex
system design updated

\section{Case Study}

Applied the system to Shenzhen's dataset. We demonstrate the efficiency.

\subsection{OVERVIEW Case 1: City Folding}

compare the distorted maps from various groups

Different different groups of people and see their patterns across the city.

Group 1 is defined as xxxxx. The popular visiting places

\subsection{DETAIL Case 2: Look into Detail}

check the detail information of a group

To compare the Group 1 and Group 2, the similar places and different places...