### 2019-04-20 answer.tex
revision begin

%%%%%%%%%%%%%%%%%%%%%%% answer list.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% For answering the review comments of JOV.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%

\documentclass[UTF8]{article}
\usepackage{geometry}
\geometry{left=1.0cm}

\usepackage{epstopdf}
\usepackage{mathptmx}
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{verbatim}

\usepackage{color}
\newcommand{\m}{\textcolor{red}}

\usepackage{array}
\usepackage{multirow}
\usepackage{tabularx}

\usepackage{algpseudocode}
\usepackage{natbib}
\usepackage[misc]{ifsym}

\begin{document}

\begin{table}[h]
\centering
\caption{focus on}
    \begin{tabular}{|p{4cm}|p{9cm}|p{4cm}|}
    \hline
    Question & Details & Answer\\
    \hline

    Design Rationale & (R1) The techniques presented in this paper, the glyph visualization, the Tsne, and the 2.5D spatial view, are not novel. They are inappropriately combined without any design rationale or considerations.\newline & A:\color{red}{We performed a in-depth interview with domain exports to understand their demands (added as a new Subsection XX), and refine the system design by several iterations. The visual design is reasoned according to the demands and derived rationales (updated in Section XX}\\

    Design Rationale & (R2) It is unclear how these tasks were obtained. It is unclear who will be the end-users. These are missing links between the tasks and the design considerations. These tasks and design considerations are unlinked to the design of the system.\newline  & A:\color{red}{Adding experts discussion to rewrite these parts.}\\

    The glyph design & (R1) The cartoon-style glyph visualization is particularly inefficiently designed. I had hard time to use this view to differentiate profile attributes for a moderate number of people. Basically it was too distracted and difficult to recall and concentrate. I would strongly suggest using simpler glyph designs or even textual labels to represent and compare multi-attributes for groups of targets.\newline & A:\color{red}{We took reviewer' comment into consideration, and added a star-like glyph designed (Section XX), to compensate the understanding of individuals' attributes from the abstract view.}\\

    The TSNE & (R1) The Tsne view is a routine for multidimensional exploration and seems useful, but the usefulness is never mentioned in the case studies.\newline & A:\color{red}{We clarified the usage of T-sne projection in Case 1, where T-sne helps to find groups with interested high dimensional attributes.}\\

    2.5D view & (R1) The 2.5D spatial view is neither effective. It does not make sense to directly reuse the 2.5D-design in this application domain. (R2)The 2.5D design for visualizing geo data is really ineffective.\newline & A:\color{red}{We took reivewer' advices of ineffective 2.5D and changed the spatial view to conventional 2D view.}\\

    Case studies to demonstrate its usefulness & (R1)I hardly found any interesting points or unexpected findings from the studies. Many conclusions have been studied and discussed in previous visual analytics papers  \newline & A:\color{red}{We updated the case studies, including the observations in ..., ....}\\

    Case studies & (R1) There even lacks a brief description about any ``individual" characteristics. How would you explore and compare the mobility- characteristics-correlation at the individual level (as claimed as a major contribution of this paper)?\newline & A:\color{red}{We enriched the exploration in ``individual" movement by... (Section XX).}\\

    User Study & (R2) I don't how the system will be used in practice. The authors make this clear and conduct interview studies with the users to further demonstrate the effectiveness of the visualization. & A:\color{red}{We performed an interview with domain experts in usage, and reported their feedback in Section XXX.}\\

    \hline
    \end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{not so important}
    \begin{tabular}{|p{3cm}|p{9cm}|p{5cm}|}
    \hline
    Question & Details & Answer \\
    \hline
    Writing & (R1)The paper is not well written. Typos, awkward sentences, and grammatical errors are throughout the paper.\newline & A: \\

    Related work & (R1)The related work is insufficiently discussed and is not well organized.\newline & A: \\

    \hline
    \end{tabular}
\end{table}

\end{document}
% end of file template.tex